### 2015-07-14 writeup.tex
add latex.

\documentclass[11pt]{article}

\begin{document}

This is my first LaTeX document.

\end{document}