### 2015-07-20 A3/latex/writeup.tex
Add command command execution and beggining of shell string interpretation.

\documentclass[11pt]{article}

\begin{document}

\section{Design}

\section{Worklog}

\section{Challenges}

\section{Questions}
\textbf{Point of assignment}: To go over how the command shell works. To teach how

\textbf{Correctness testing}:

\textbf{Learn}:

\end{document}