### 2015-08-14 A5/writeup.tex
last minute commit.

\documentclass[11pt]{article}

\sepackage{graphicx}

\begin{document}
Elliott Capek \\

\section{Design}

\section{Worklog}

\section{Challenges}

\section{Questions}
\textbf{Point of assignment}:

\textbf{Correctness testing}:

\textbf{What was learned}:

\end{document}