### 2015-10-15 comp_model/latex/a1.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Computational Modeling HW 1}
\author{Elliott Capek}

\begin{document}

\maketitle{}

\section{1.4: Trapezoidal method coefficients}
\begin{equation}
  \sum_{j=0}^r \alpha_jV_{n+j} = h \sum_{j=0}^r \beta_j f(t_{n+j}, V_{n+j})\\
\end{equation}

The trapezoidal method is this:\\
\begin{equation}
  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
\end{equation}

By examining these two equations, we can see that:

\begin{align}
  \alpha_0 = -1 &\hspace{2cm} \beta_0 = \frac{1}{2}\\
  \alpha_1 = 0 &\hspace{2cm} \beta_1 = 1\\
  \alpha_2 = 0 &\hspace{2cm} \beta_2 = 1\\
  ...  &\hspace{2cm} ...\\
  \alpha_{r-1} = 0 &\hspace{2cm} \beta_{r-1} = 1\\
  \alpha_r = 1 &\hspace{2cm} \beta_1 = \frac{1}{2}\\    
\end{align}

\begin{equation}
  V_{n+1} = V_n + h\Big( \frac{1}{2}f(V_n, t_n) + \frac{1}{2}f(V_{n+1}, t_{n+1}) \Big)
\end{equation}

If we let our function f take the following form: $f(x, t) = \alpha x$ where $x(0) = x_0$, we can see that:

\begin{align}
  V_{n+1} &= V_n + \frac{h}{2}\Big( f(V_n, t_n) + f(V_{n+1}, t_{n+1}) \Big)\\
  V_{n+1} &= V_n + \frac{h}{2}\Big( \alpha V_n + \alpha V_{n+1} \Big)\\
  V_{n+1} &= V_n + \frac{\alpha hV_n}{2} + \frac{\alpha hV_{n+1}}{2}\\
  V_{n+1} \Big(1 - \frac{\alpha h}{2}\Big) &= V_n \Big(1 + \frac{\alpha h}{2}\Big) \\
  V_{n+1} &= V_n \frac{1 + \frac{\alpha h}{2}}{1 - \frac{\alpha h}{2}} \\
  V_{n+1} &= V_n \frac{2 + \alpha h}{2 - \alpha h} \\
  V_{n+1} &= V_0 \Big(\frac{2 + \alpha h}{2 - \alpha h}\Big)^{n+1} \\    
\end{align}

From this we can see that as long as the following condition is satisfied, we will have a converging solution:

\begin{align}
  |2 + \alpha h| \leq |2 - \alpha h|\\
\end{align}
\section{1.9: Order of Accuracy for Trapezoidal and Euler Methods}
We want to analytically represent the error terms for the Euler and Trapezoidal methods and show how they depend on our timestep, $h$.\\

First we solve for the behavior of our function using a Taylor Series. Let $x(t)$ be our function and $v(t) = \frac{dx(t)}{dt}$. Let $N = \frac{t}{h}$, where t is the current time and h the timestep. Finally, let $x(t_0) = 0$.\\

\begin{align}
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  v(t) &= v(t_i) + (t-t_i)v'(t_i) + (t-t_i)^2v''(t_i) + ...\\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx \int_{t_i}^{t_{i+1}} v(t_i) + (t-t_i)v'(t_i) + \frac{1}{2}(t-t_i)^2v''(t_i)  dt\\
\end{align}

We now do a u-substitution, where $S = \frac{t - t_i}{t_{i+1} - t_i}$ and $h = t_{i+1} - t_i$:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h\int_i^1 v(t_i) + hSv'(t_i) + \frac{1}{2}h^2S^2v''(t_i) dS\\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i)S + \frac{hS^2}{2}v'(t_i) + \frac{h^2S^3}{6}v''(t_i) \big) |_0^1 \\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx h \big( v(t_i) + \frac{h}{2}v'(t_i) + \frac{h^2}{6}v''(t_i) \big) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &\approx hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
\end{align}

This gives us a fairly exact value for $\int_{t_i}^{t_{i+1}} v(t)dt$. We now need to compare this with the approximations that our two methods make. For Euler's method, $\int_{t_i}^{t_{i+1}} v(t)dt = hf(t_i)$. For the Trapezoidal method, $\int_{t_i}^{t_{i+1}} v(t)dt = \frac{h}{2}\Big( f(t_{i+1}) + f(t_{i}) \Big)$. We will first plug in the Euler's method approximation to the above expression to find its error term.

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
  hv(t_i) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i)\\
  -\frac{h}{3}v''(t_i) &= v'(t_i)\\  
\end{align}

We now substitute this back into our original result:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}(-\frac{h}{3}v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{4}(v''(t_i)) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
\end{align}

As we can see, the error for the forward Euler's method is $-\frac{h^3}{12}(v''(t_i))$ per timestep. We now plug this result into our summed equation to find the total error for all timesteps:

\begin{align}
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} hv(t_i) + -\frac{h^3}{12}(v''(t_i))\\  
\end{align}

We only care about the error term, so we will focus on that. We substitute $N$ with $\frac{t - t_0}{h}$ and replace $v''(t_i)$ with a term that represents the average value of $v''(t_i)$ over the whole sum: $v''_{ave}$.

\begin{align}
  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
  E &= -N\frac{h^3}{12}v''_{ave}\\
  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\    
\end{align}

We do the same thing for the Trapezoidal method. We first find the error term:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i)\\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
  \frac{h}{2}\Big(v(t_{i+1}) + v(t_i)\Big) &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{4}v''(t_i)\\
  \frac{h}{2}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h^3}{4}v''(t_i) &= \frac{h^2}{2}v'(t_i)\\
  \frac{1}{h}\Big(-v(t_{i+1}) + v(t_i)\Big) - \frac{h}{2}v''(t_i) &= v'(t_i)\\
\end{align}

We then substitute this back into our original result:

\begin{align}
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}v'(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= hv(t_i) + \frac{h^2}{2}\Big(\frac{1}{h}\Big(v(t_{i+1}) - v(t_i)\Big) - \frac{h}{4}v''(t_i) \Big) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{4}v''(t_i) + \frac{h^3}{6}v''(t_i) \\
  \int_{t_i}^{t_{i+1}} v(t)dt &= \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\    
\end{align}

From this we can see that the error term for the trapezoidal rule is $-\frac{h^3}{12}v''(t_i)$ per timestep. We then sum over all timesteps as above:

\begin{align}
    x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} v(t)dt\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) - \frac{h^3}{12}v''(t_i)\\
  x(t) &= \sum_{i=0}^{i=N-1} \int_{t_i}^{t_{i+1}} x'(t)dt = \sum_{i=0}^{i=N-1} \frac{h}{2}\big(v(t_{i+1}) + v(t_i)\big) + \sum_{i=0}^{i=N-1} -\frac{h^3}{12}v''(t_i)\\
\end{align}

We finally solve for the error:

\begin{align}
  E &= \sum_{i=0}^{i=N-1} -\frac{h^3}{12}(v''(t_i))\\
  E &= -N\frac{h^3}{12}v''_{ave}\\
  E &= -\frac{t - t_0}{h} \frac{h^3}{12}v''_{ave}\\
  E &= -(t - t_0)\frac{h^2}{12}v''_{ave}\\
\end{align}

From this we can see that the Trapezoidal method's error is proportional to $h^2$.

\end{document}