### 2015-11-02 comp_model/latex/a2.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Computational Modeling HW 1}
\author{Elliott Capek}

\begin{document}

\maketitle{}

\section{3.2: Steady-state Population Model}

We will let $u_j < 1$ for all $j$. We also let the following be true:

\begin{equation*}
  \sum_j \beta_j \Big(\prod_{k<j}(1-u_k) \Big) = 1\\
\end{equation*}

Consider the following matrix:

\begin{equation}
  P = 
   \begin{pmatrix}
     \beta_1 & \beta_2 & ... & \beta_{N-1} & \beta_N\\
     1-u_1 & & & &\\
     & 1-u_2 & & &\\
     & & ... & &\\
     & & 1-u_{N-1} & &\\
     & & & 1-u_{N} & 0\\
   \end{pmatrix}
\end{equation}
\end{document}