### 2015-11-23 classical/latex/a8.tex
Add stuffs.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Classical Mechanics Assignment 8}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: }
\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}

\section{Problem Two: }
\textbf{The potential for an non-harmonic oscillator is:}

\begin{equation*}
  U(x) = \frac{kx^2}{2} + \frac{bx^4}{4}\\
\end{equation*}

\textbf{where k and b are positive constants. Find Hamilton’s equations of motion. No final discussion needed.}

\section{Problem Three: }
\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal). Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}

\section{Problem Four: }
\textbf{A thin hoop of radius R and mass M oscillates in its own plane hanging from a single fixed point. Attached to the hoop is a small mass M constrained to move (in a frictionless manner) along the hoop. The inertia of the loop for rotation around a point on the loop is $I=2MR^2$. Consider only small oscillations, and show that the eigenfrequencies are:}

\begin{equation*}
  \omega_1 = \sqrt{\frac{2g}{R}}
\end{equation*}

\begin{equation*}
  \omega_2 = \sqrt{\frac{g}{2R}}
\end{equation*}

\end{document}