### 2015-11-28 comp_model/latex/der.tex
stuff

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage{amsmath}
\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Free energy derivative}
\author{Elliott Capek}

\begin{document}
\maketitle{}

\section{Introduction}
Doing the derivation of free energy:

\begin{align*}
  G &= V(\phi(r)) + \frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}
\end{align*}

\begin{align*}
  \frac{dG}{d\phi} &= \frac{d}{d\phi}V(\phi(r)) + \frac{d}{d\phi}\frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}\\
  &= -\phi(1-\phi)(\phi-\frac{1}{2}+n) + \frac{d}{d\phi}\frac{\delta^2\left(\nabla\phi(r)\right)^2}{2}\\
\end{align*}

\end{document}