### 2016-01-20 quantum/latex/a2.tex
Add changes from last nite

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
% \usepackage{physics}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{9.4: Forbidden region probability for QHO}
For this problem we want to calculate the probability of measuring a QHO ground-state object outside
of the classically allowed region. We do this by calculating the turning points of a classical HO,
ie finding the x-values where energy equals potential, and then integrating $\psi^*\psi$ over the
forbidden region to find the chance of forbidden observation.\\

We first find the turning point for the ground state $n=0$:
\begin{align*}
  V_{HO} &= E_0\\
  \frac{1}{2}m\omega^2x^2 &= \hbar\omega\left(0+\frac{1}{2}\right)\\
  x_0 &= \sqrt{\frac{\hbar}{m\omega}}\\
\end{align*}

We then find the probability of finding the particle in this region:

\begin{align*}
  P_{forb} &= \int_{\infty}^{-x_0} \psi_0^*\psi_0 dx + \int_{x_0}^{\infty} \psi_0^*\psi_0 dx\\
  &= \int_{\infty}^{-x_0} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}dx
  + \int_{x_0}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar} dx\\
  &= \sqrt{\frac{m\omega}{\pi\hbar}} \left(\frac{-\hbar}{m\omega}\right)\left(
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{-\infty}^{-\sqrt{\frac{\hbar}{m\omega}}} +
    \frac{1}{x}e^{-m\omega x^2 / \hbar} \Bigg|_{\sqrt{\frac{\hbar}{m\omega}}}^{\infty}\right)\\
  &= \frac{2}{\sqrt{\pi}} = 0.41\\
\end{align*}

%% \section{9.9: QHO expectation values via integration, operator method}
%% In this problem we use both wave function integration and the operator method to find the
%% expectation values of various observables: x, p, $x^2$, $p^2$.

%% \textbf{a.) Ground state wave function calculations}
%% \begin{align*}
%%   <x> &= \int_{-\infty}^{\infty} \psi^*x\psi dx\\
%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x dx = 0\\
%% \end{align*}
%% We don't need to perform the $<x>$ integral because it involves integrating an odd function over
%% an even domain, which always results in zero. This is to be expected, because the wave ground-state
%% wave function is symmetric across zero.\\

%% \begin{align*}
%%   <p> &= \int_{-\infty}^{\infty} \psi^* -i\hbar \frac{d}{dx}\psi dx\\
%%   &= \int_{-\infty}^{\infty}
%%   -i\sqrt{\frac{m\omega}{\pi\hbar}}(-m\omega x)e^{\frac{-m\omega x^2}{2\hbar}}dx = 0\\
%% \end{align*}
%% This integral is zero since we are again integrating an odd function over an even domain. This
%% makes sense because, since the spacial distribution of the wave function is equal on both sides
%% of the x-axis, the velocities (and hence momenta) should be equal as well.\\

%% \begin{align*}
%%   <x^2> &= \int_{-\infty}^{\infty} \psi^*x^2\psi dx\\
%%   &= \int_{-\infty}^{\infty} \sqrt{\frac{m\omega}{\pi\hbar}}e^{-m\omega x^2 / \hbar}x^2 dx\\
%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}}\int_{0}^{\infty} e^{-m\omega x^2 / \hbar}x^2 dx\\
%%   &= 2\sqrt{\frac{m\omega}{\pi\hbar}} *
%%   \frac{\sqrt{\pi}}{4(\frac{m\omega}{\hbar})^{3/2}} \hspace{2cm}\mbox{Wolfram}\\
%%   &= \frac{\hbar}{2m\omega}\\
%% \end{align*}
%% Here we exploit the fact that integrating an even function over an even domain is equivalent to
%% doubling the integral of the function over half the domain. The answer makes sense, since increasing
%% $m\omega$ corresponds to making the quadratic potential steeper, hence making the wave function
%% tighter.\\

%% \begin{align*}
%%   <p^2> &= \int_{-\infty}^{\infty} \psi^* \left(-\hbar^2 \frac{d^2}{dx^2}\right)\psi dx\\
%%   &= \sqrt{\frac{m\omega}{\pi\hbar}}\int_{-\infty}^{\infty} -\hbar^2e^{\frac{-m\omega x^2}{\hbar}}
%%   \left(\frac{m^2\omega^2x^2}{\hbar^2} - \frac{m\omega}{\hbar}\right)dx\\
%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}\left(
%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m^2\omega^2x^2}{\hbar^2}dx -
%%   \int_{-\infty}^{\infty} e^{\frac{-m\omega x^2}{\hbar}}\frac{m\omega}{\hbar}\right)dx\\
%%   &= -\hbar^2\sqrt{\frac{m\omega}{\pi\hbar}}
%%   \left(\frac{1}{2}\sqrt{\frac{m\pi\omega}{\hbar}} -
%%   \sqrt{\frac{m\pi\omega}{\hbar}}\right) \hspace{2cm}\mbox{Wolfram}\\
%%   &= \frac{\hbar m\omega}{2}\\
%% \end{align*}
%% This answer is intuitive, since increasing mass as well as oscillatory speed should both increase
%% the expected momentum.

%% \textbf{b.) Any-state operator method calculations}
%% Here we define the position and momentum operators in terms of the raising and lowering operators:
%% \begin{align*}
%%   \hat{x} &= \sqrt{\frac{\hbar}{2m\omega}}\left(a + a^\dagger\right)\\
%%   \hat{p} &= i\sqrt{\frac{m\omega\hbar}{2}}\left(a^\dagger - a\right)\\
%% \end{align*}
%% We then solve for the various expectation values using these definitions:

%% \begin{align*}
%%   <x> &= \bra{n}x\ket{n}\\
%%   &= \sqrt{\frac{\hbar}{2m\omega}}\bra{n}\left(a + a^\dagger\right)\ket{n}\\
%%   &= \sqrt{\frac{\hbar}{2m\omega}}
%%   \left(\sqrt{n}\bra{n}\ket{n-1} + \sqrt{n+1}\bra{n}\ket{n+1}\right) = 0\\
%% \end{align*}
%% This answer indicates that all pure QHO wave functions are centered around zero. This makes
%% sense since the QHO wave functions are all made up of either even or odd functions. When squared,
%% as is done when finding the probability, the equation becomes entirely even, meaning it is
%% symmetric across the y-axis. From this we can see that all QHO wave functions have an expectation
%% value of zero for x-measurement.

%% \begin{align*}
%%   <p> &= \bra{n}\hat{p}\ket{n}\\
%%   &= i\sqrt{\frac{m\hbar\omega}{2}}\bra{n}\left(a^\dagger-a\right)\ket{n}\\
%%   &= i\sqrt{\frac{m\hbar\omega}{2}}
%%   \left(\sqrt{n+1}\bra{n}\ket{n+1} - \sqrt{n}\bra{n}\ket{n-1}\right) = 0\\
%% \end{align*}
%% This is an interesting result. It is intuitive that any wave function should have an expectation of
%% zero for momentum, else the wave function would move through space due to a nonzero average
%% velocity.

%% \begin{align*}
%%   <x^2> &= \bra{n}x^2\ket{n}\\
%%   &= \frac{\hbar}{2m\omega}\bra{n}\left(a + a^\dagger\right)\left(a + a^\dagger\right)\ket{n}\\
%%   &= \frac{\hbar}{2m\omega}\bra{n}
%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
%%   &= \frac{\hbar}{2m\omega}
%%   \left(aa + aa^\dagger + a^\dagger a + a^\dagger a^\dagger\right)\ket{n}\\
%%   &= \frac{\hbar}{2m\omega} \left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2} + (n+1)\bra{n}\ket{n} +
%%   (n)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
%%   &= \frac{\hbar}{2m\omega} \left(2n+1\right)\\
%% \end{align*}
%% This is an interesting result which indicates that for each extra rung on the ladder, the
%% expectation of the squared position increases linearly, meaning the expectation of the magnitude
%% of x increases as $\sqrt{n}$. This makes sense, since increasing energy levels have less and less
%% room to expand in the quadratic potential well. The $n=0$ state corresponds to the wave function
%% solution for the ground state found above.\\

%% \begin{align*}
%%   <p^2> &= \bra{n}p^2\ket{n}\\
%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(a^\dagger-a\right)\left(a^\dagger-a\right)\ket{n}\\
%%   &= -\frac{m\hbar\omega}{2}\bra{n}
%%   \left(a^\dagger a^\dagger - a^\dagger a - aa^\dagger + aa\right)\ket{n}\\
%%   &= -\frac{m\hbar\omega}{2}\bra{n}\left(\sqrt{n-1}\sqrt{n-2}\bra{n}\ket{n-2}
%%   - (n)\bra{n}\ket{n} - (n+1)\bra{n}\ket{n} + \sqrt{n+1}\sqrt{n+2}\bra{n}\ket{n+2}\right)\\
%%   &= \frac{\hbar m\omega}{2}\left(2n + 1\right)
%% \end{align*}
%% This result is similar to the above one for $x^2$ with different units. It is still a $\sqrt{n}$
%% relationship for momentum, which makes sense.\\

%% \textbf{c.) Uncertainty principle}\\
%% Since the above two cases agree with each other, we check that the more general case of part
%% \textbf{b.} agrees with the uncertainty principle.

%% \begin{align*}
%%   \Delta x &= \sqrt{<x^2>-<x>^2}\\
%%   &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}\\
%%   \Delta p &= \sqrt{<p^2>-<p>^2}\\
%%   &= \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
%%   \Delta x \Delta p &= \sqrt{\frac{\hbar}{2m\omega} \left(2n+1\right)}
%%   \sqrt{\frac{\hbar m\omega}{2}\left(2n + 1\right)}\\
%%   &= \frac{\hbar}{2}(2n+1)\\
%% \end{align*}
%% This is an interesting result which indicates that uncertainty increases as the pure wave function
%% gets more energetic. Ground state is the state with maximum certainty. The higher states get
%% progressively more uncertain, which makes sense since both the velocity and the well get larger
%% for higher states.\\

%% \section{9.14: Unknown QHO wavefunction}
%% We are given the wave function $\psi(x,0) = A\left[1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%   2\frac{m\omega}{\hbar}x^2\right]$. Because its highest-order polynomial term is $x^2$, we know
%% that it must be made up of the $n = 0, 1, 2$ energy states. We now figure out the coefficients for
%% each:

%% \begin{align*}
%%   c_0 &= \int_{-\infty}^{\infty} \phi_0^*\psi dx\\
%%   &= A\int_{-\infty}^{\infty} \left(\frac{m\omega}{\pi\hbar}\right)^{1/2}e^{-m\omega x^2/\hbar}
%%   \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%   2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/(2\hbar)}dx\\
%%   &= 2\\
%%   c_1 &= A\int_{-\infty}^{\infty} \frac{2\sqrt{\frac{m\omega}{\hbar}x}}{\sqrt{2}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= -\frac{3}{\sqrt{2}}\\
%%  c_2 &= A\int_{-\infty}^{\infty} \frac{\frac{4m\omega}{\hbar}x^2 - 2}{\sqrt{8}}
%%     \left(\frac{m\omega}{\pi\hbar}\right)^{1/2} \left(1 - 3\sqrt{\frac{m\omega}{\hbar}}x +
%%     2\frac{m\omega}{\hbar}x^2\right)e^{-m\omega x^2/\hbar}dx\\
%%  &= \sqrt{2}\\
%% \end{align*}

%% From this we can see that $\psi(x,0) = \sqrt{\frac{21}{2}}
%% \left(2\ket{0} + -\frac{3}{\sqrt{2}}\ket{1} + \sqrt{2}\ket{2}\right)$. It is then easy to compute
%% the expectation value: \\

%% \begin{align*}
%%   <E> &= \sigma \hbar\omega\left(n+\frac{1}{2}\right)|c_n|^2\\
%%   &= \frac{2 + \frac{27}{4} + 5}{\frac{21}{2}} = 1.3 eV\\
%% \end{align*}

\section{6: Position, momentum commutator}
Here we find the value of the position, momentum commutator:
\begin{align*}
  [\hat{x}, \hat{p}] &= \hat{x}\hat{p} - \hat{p}\hat{x}\\
  &= \left(\hat{x}\hat{p}\psi(x) - \hat{p}\hat{x}\psi(x)\right)\frac{1}{\psi(x)}\\
  &= \left(\hat{x}\left(-i\hbar\frac{d}{dx}\right)\psi(x) - \hat{p}x\psi(x)\right)\frac{1}{\psi(x)}\\
  &= \left(-i\hbar x \frac{d\psi}{dx} + i\hbar\left(\psi(x)
  + x\frac{d\psi}{dx}\right)\right)\frac{1}{\psi(x)}\\
  &= \left(i\hbar\psi(x)\right)\frac{1}{\psi(x)}\\
  &= i\hbar\\
\end{align*}
Here we use the definitions of the position and momentum operators to find the value of their
commutator. We need to operate on a generic wave function in order to do algebra on the
operators, since the differential in the momentum operator is meaningless without a function
to act on. This result is critical, since it is at the heart of the Heisenberg Uncertainty
Principle: position and momentum don't commute, so they do not share a basis and they are not
co-measurable. This is the mathematical representation of the popular saying ``you cannot know
the position and momentum of a particle simultaneously.''

\section{7: Commutator algebra}

\begin{align*}
  \left[\hat{A},\hat{A}^n\right] &= AA^n - A^nA = A^{n+1} - A^{n+1} = 0\\
  \left[\hat{A},\hat{B}\right] &= \hat{A}\hat{B} - \hat{B}\hat{A}
  = -\left(\hat{B}\hat{A} - \hat{A}\hat{B}\right) = -\left[\hat{B},\hat{A}\right]\\
  \left[\hat{A},c\hat{B}\right] &= \hat{A}c\hat{B} - c\hat{B}\hat{A}
  = c\left(\hat{A}\hat{B} - \hat{B}\hat{A}\right) = c\left[\hat{A},\hat{B}\right]\\
  \left[\hat{A},\left(\hat{B}+\hat{C}\right)\right] &= \hat{A}\left(\hat{B} + \hat{C}\right)
  - \left(\hat{B} + \hat{C}\right)\hat{A} = \hat{A}\hat{B} - \hat{B}\hat{A}
  + \hat{A}\hat{C} + \hat{C}\hat{A} = \left[\hat{A},\hat{B}\right] + \left[\hat{A},\hat{C}\right]\\
  \left[\hat{A},\hat{B}\hat{C}\right] &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{C}\hat{A}
  = \hat{A}\hat{B}\hat{C} - \hat{B}\hat{A}\hat{C} + \hat{B}\hat{A}\hat{C} - \hat{B}\hat{C}\hat{A}
  = \left[\hat{A},\hat{B}\right]\hat{C} + \hat{B}\left[\hat{A},\hat{C}\right]\\
  \left[\hat{A},\left[\hat{B},\hat{C}\right]\right]
  + \left[\hat{B},\left[\hat{C},\hat{A}\right]\right]
  &+ \left[\hat{C},\left[\hat{A},\hat{B}\right]\right] =
  \left[\hat{A},\hat{B}\hat{C} - \hat{C}\hat{B}\right]
  + \left[\hat{B},\hat{C}\hat{A} - \hat{A}\hat{C}\right]
  + \left[\hat{C},\hat{A}\hat{B} - \hat{B}\hat{A}\right] = \\
  &= \hat{A}\hat{B}\hat{C} - \hat{A}\hat{C}\hat{B}
  + \hat{B}\hat{C}\hat{A} - \hat{B}\hat{A}\hat{C}
  + \hat{C}\hat{A}\hat{B} - \hat{C}\hat{B}\hat{A}\\
\end{align*}

\end{document}