### 2016-01-31 quantum/latex/a4.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Three}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{10.8: Magnetic Perturbation Spin 1}
For this problem we want to find the energy perturbations due to a magnetic field perturbation
in the x direction, while remaining in the z basis. To do this we first find the
perturbed Hamiltonian, which is just the x spin-1 matrix times a factor, and
then use the formula
%$E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}$
to find the second-order energy corrections.\\

We first find the perturbed Hamiltonian:

\begin{align*}
  H' &= \vec{\mu} \cdot \vec{B_2}\\
  &= \frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
  &= \frac{\hbar\omega_2}{2} S_x
  \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
  &=
  \begin{pmatrix}
    0 & \frac{\hbar\omega_2}{2} & 0\\
    \frac{\hbar\omega_2}{2} & 0 & \frac{\hbar\omega_2}{2}\\
    0 & \frac{\hbar\omega_2}{2} & 0\\
  \end{pmatrix}
\end{align*}

Then we just apply the second-order perturbation energy equation to find our
energy changes, noticing that there will be no first-order corrections because
$H'$ has zeroes along its diagonal.

\begin{align*}
  E_n^{(2)} = \sum_{m\noteq n}\frac{|\bra{n^{(0)}}H'\ket{n^{(0)}|^2}{(E_n^{(0)}-E_m^{(0)}}}}\\
\end{align*}

\begin{align*}
  E_{\frac{-1}{2}} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

\begin{align*}
  E_{0} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

\begin{align*}
  E_{\frac{1}{2}} &=
  \frac{|H'_{\frac{-1}{2},0}|^2}{E_{0} - E_{\frac{-1}{2}}} +
  \frac{|H'_{\frac{-1}{2},\frac{1}{2}}|^2}{E_{\frac{1}{2}} - E_{\frac{-1}{2}}}\\
  &= \frac{\hbar^2\omega_2^2}{2\hbar\omega_0}
  = \frac{\hbar\omega_2^2}{2\omega_0}\\
\end{align*}

\end{document}