### 2016-02-03 bioinformatics/latex/a2.tex
all

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Bioinformatics Assignment 2}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Scoring Alignments}
Here we are asked to compute the score for different given alignments.\\

\textbf{A.} 3\\
\textbf{B.} 1\\

\section{Problem Two: BLAST}

\section{Problem Four: Database size}
We use the definition of E-value to compute database size, using an E-value of $e^{-27}:\\

\begin{align*}
  m &= \frac{E2^{S'}}{n} = \frac{e^{-27}2^{121}}{150} = 3.3*10^{22}\\
\end{align*}

\section{Problem Seven: GC Content}
GC content is conserved when taking reverse complement, since Gs and Cs map to
each other.\\

For original probability $p_{N0}$ and reverse complement probability $P_N'$:\\

\begin{align*}
  p_{G0} &= p_C'\\
  p_{C0} &= p_G'\\
  p_G &= \frac{p_{G0} + p_G'}{2} = \frac{p_{G0} + p_{C0}}{2} = \frac{0.63}{2} = 0.315\\
  p_C &= 0.315\\
  p_A &= 0.185\\
  p_T &= 0.185\\
\end{align*}

\section{Problem Eight: MEME}
Here we enter some MEME commands to find motifs:\\
\textbf{A.} meme proteins.fasta -protein -mod zoops -minw 10\\
\textbf{B.} meme sequences.fa -rna -mod anr -minw 5 -maxw 10\\
\textbf{C.} meme cores.fasta -dna -mod oops -nmotifs 5 \# -revcomp not specified,
only search forward strand\\

\end{document}