### 2016-02-24 quantum/latex/a7.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{marvosym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Quantum Mechanics Assignment Seven}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 1: Relativistic Fine Correction}

This problem asks us to work through parts of the derivation of perturbation theory applied to
the relativistic correction to the Hydrogen energy spectrum.\\

\textbf{a.)}Here we show how to find the expectation value of $p^4$ more easily:\\

\begin{align*}
  &\bra{n,\ell,m}p^4\ket{n,\ell,m}\\
  &= \bra{n,\ell,m}p^2p^2\ket{n,\ell,m}\\
  &= 4m^2\bra{n,\ell,m}\left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)
  \left(H_0 + \frac{e^2}{4\pi\epsilon r}\right)\ket{n,\ell,m}\\
  &= 4m^2\left(\bra{n,\ell,m}H_0^2\ket{n,\ell,m}
  + \bra{n,\ell,m}H_0\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}H_0\ket{n,\ell,m}
  + \bra{n,\ell,m}\frac{e^4}{16\pi^2\epsilon^2r^2}\ket{n,\ell,m}\right)\\
  &= 4m^2\left((E_n^{(0)})^2
  + E_n^{(0)}\bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}
  + \bra{n,\ell,m}\frac{e^2}{4\pi\epsilon r}\ket{n,\ell,m}E_n^{(0)}
  + \frac{e^4}{16\pi^2\epsilon^2r^2}\bra{n,\ell,m}\frac{1}{r^2}\ket{n,\ell,m}\right)\\
  &= 4m^2\left((E_n^{(0)})^2
  + \frac{e^2E_n^{(0)}}{4\pi\epsilon}\braket{\frac{1}{r}}
  + \frac{e^2}{4\pi\epsilon}\braket{\frac{1}{r}}E_n^{(0)}
  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
  &= 4m^2\left((E_n^{(0)})^2
  + \frac{e^2E_n^{(0)}}{2\pi\epsilon}\braket{\frac{1}{r}}
  + \frac{e^4}{16\pi^2\epsilon^2r^2}\braket{\frac{1}{r^2}}\right)\\
\end{align*}

\textbf{b.)}\\

\textbf{c.)} We perform the following integrals using Mathematica:

\begin{align*}
  \braket{\frac{1}{r}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
  4r\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z}{4a}\\
  \braket{\frac{1}{r^2}}_{20} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
  4\left(1-\frac{Zr}{2a}\right)^2e^{-Zr/a}dr = \frac{Z^2}{4a^2}\\
  \braket{\frac{1}{r}}_{21} &= \left(\frac{Z}{2a}\right)^3\int_0^\infty
  \frac{Z^2r}{3a^2}e^{Zr/a}dr = \frac{Z}{4a}\\
  \braket{\frac{1}{r^2}}_{21} &= \left(\frac{Z^5}{24a^5}\right)^3\int_0^\infty
  r^2e^{-Zr/a}dr = \frac{Z^2}{12a^2}\\
\end{align*}

These are the desired results. We then assume this is true for all n, $\ell$ quantum numbers.\\

\textbf{d.)}
Here we expand McIntyre (12.26) and show that it is equivalent to (12.30):\\

\begin{align*}
  \braket{H'_{rel}} &=
  -\frac{1}{2mc^2}\left((E_n^0)^2 +
  2\frac{e^2}{4\pi\epsilon_0}E_n^0\braket{\frac{1}{r}} +
  \frac{e^4}{16\pi^2\epsilon_0^2}\braket{\frac{1}{r^2}}\right)\\
  &= -\frac{1}{2mc^2}\left(\frac{\alpha^4m^2c^4}{4n^4}
  + 2\frac{e^2}{4\pi\epsilon_0}\left(\frac{-\alpha^2mc^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{e^4}{16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2}\right)\right)\\
  &= \frac{-\alpha^4mc^2}{8n^4}
  + \frac{e^2}{4\pi\epsilon_0}\left(\frac{\alpha^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{-e^4}{2*16\pi^2\epsilon_0^2}\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
  &= \frac{-\alpha^4mc^2}{8n^4}
  + a_0\alpha^2mc^2\left(\frac{\alpha^2}{2n^2}\right)
  \left(\frac{1}{n^2a_0}\right) +
  \frac{-1}{2}a_0^2\alpha^4m^2c^4\left(\frac{1}{(\ell+1/2)n^3a_0^2mc^2}\right)\\
  &= \alpha^4mc^2\left(\frac{3}{8n^4}\right) +
  \frac{1}{2}\alpha^4mc^2\left(\frac{-1}{(\ell+1/2)n^3}\right)\\
  &= -\frac{1}{2}\alpha^4mc^2\left(\frac{1}{n^3\left(\ell+\frac{1}{2}\right)} -
    \frac{3}{4n^4}\right)\\
\end{align*}

This was an interesting problem. It was good to go over how the relativistic fine-
structure corrections are derived, especially how to deal with a $1/\hat{r}$
operator. At first I was confused about how to deal with an inverse operator, but
then I realized that it was just a normal operator, like $\hat{r^2}$. Exploiting
the fact that the complex conjugate Hamiltonian acts on bras the same way that the
normal Hamiltonian acts on kets was another important method to solving this problem.\\

\section{2: Spectroscopy Numbers}
Here we discuss the spectroscopic labels and splitting we see in the figure in the
problem.\\

The labels are kind of like spectroscopic notation. The upper term (3d, 2p, etc...) is
the n and $\ell$ number for the state, where $\ell$ is translated into alphabet
notation. The subscript is the F-number (L+S) for that orbital. The spin of the electron can be
$\pm\frac{1}{2}$, so the F-number plus or minus $\frac{1}{2}$ is the $\ell$-number. $3d_{5/2}$ is
the n=3, $\ell$=2 orbital \textit{where the angular momentum and spin momentum combine}. There is
another orbital where they are subtracted.\\

The $3p_{3/2}$ and $3d_{3/2}$ orbitals both have corrections of 1.62GHz, as seen in the table. They
are made up of the following linear combinations of uncoupled states:\\

\begin{align*}
  \ket{n,F,M_F} &= \ket{n,M_I,M_s}\\
  \ket{3,\frac{3}{2}, \frac{ 3}{2}} &= \ket{3,1,\frac{1}{2}}\\
  \ket{3,\frac{3}{2}, \frac{ 1}{2}} &= \frac{1}{\sqrt{3}}\ket{3,1,\frac{1}{2}}
  + \sqrt{\frac{2}{3}}\ket{2,0,\frac{1}{2}}\\
  \ket{3,\frac{3}{2}, \frac{-1}{2}} &= \sqrt{\frac{2}{3}}\ket{3,0,-\frac{1}{2}}
  + \frac{1}{\sqrt{3}}\ket{3,-1,\frac{1}{2}}\\
  \ket{3,\frac{3}{2}, \frac{-3}{2}} &= \ket{3,-1,-\frac{1}{2}}\\
\end{align*}

I liked this problem. I didn't get to finish it, but it was interesting thinking about how to use
Clebsch-Gordan tables to compare bases. It was also helpful thinking about how there are different
quantum states for when angular momenta combine vs subtract from each other. This doesn't make
sense but is an important part of quantum mechanics.\\

\section{3: Energies}
Here we calculate many different energies and show how they are related.\\

\textbf{a.) Unperturbed energies}
\begin{align*}
  E_1^0 - E_2^0 &= -13.7eV\left(\frac{1}{4} - \frac{1}{1}\right)\\
  &= 10.275eV\\
  &= \frac{3}{8}\alpha^2mc^2\\
  &= 2.48 * 10^6 GHz\\
\end{align*}

\textbf{b.) SO corrections to energy}
\begin{align*}
  E_{S}' &= \frac{1}{4}\alpha^4mc^2\frac{j(j+1)
    - \ell(\ell+1) - \frac{3}{4}}{n^3\ell(\ell+\frac{1}{2})(\ell+1)}\\
  E_{n=2,\ell=0} &= 0\\
  E_{n=2,\ell=1} &= \frac{-29}{384}\alpha^4mc^2\\
  &= -1.1 *10^{-4} eV\\
  &= 0.026 GHz\\
  E_{n=1,\ell=0} &= 0\\
\end{align*}

\textbf{c.) Relativistic corrections to energy}
\begin{align*}
  E_{rel}' &= -\frac{1}{2}\alpha^4mc^2
  \left(\frac{1}{n^3(\ell+\frac{1}{2})} - \frac{3}{4n^4}\right)\\
  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2\\
  &= -2.6*10^{-5} eV\\
  &= 6.39 GHz\\
  E_{n=2,\ell=0} &= -\frac{13}{128}\alpha^4mc^2\\
  &= 1.4*10^{-4} eV\\
  &= -35.6 GHz\\
  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2\\
  &= 9.1*10^{-4} eV\\
  &= -219.2 GHz\\
\end{align*}

\textbf{d.) Total fine-structure corrections}

\begin{align*}
  E_{n=2,\ell=1} &= -\frac{7}{384}\alpha^4mc^2 = -2.6*10^{-5} eV = -6.39 GHz\\
  E_{n=2,\ell=0} &= -\frac{17}{96}\alpha^4mc^2 = -2.5*10^{-4} = -35.6 GHz\\
  E_{n=1,\ell=0} &= -\frac{5}{8}\alpha^4mc^2 = -9.1*10^{-4} eV = -219.2 GHz\\
\end{align*}

\textbf{e.) }
\begin{align*}
  E_{n=2,\ell=1}: 6.39 GHz \rightarrow \lambda = 4.69*10^7 nm = .0496m\\
  E_{n=2,\ell=0}: 35.6 GHz \rightarrow \lambda = 8.42*10^6 nm = 0.00842m\\
  E_{n=1,\ell=0}: 219.2 GHz \rightarrow \lambda = 1.37*10^6 nm = 0.0137m\\
\end{align*}

\textbf{f.) Rest mass}
The reduced mass must be used! The electrons are travelling at roughly 1\% the speed
of light, so they feel relativistic effects non-negligibly.\\

This was a cool, long problem! It is interesting that the $\ell=0$ states have no
SO corrections, since they have no angular momentum and hence there is no apparent
B-field that they see due to ``orbiting'' the proton. The numbers are weird, it is odd
to me that there are such large differences (nearly a factor of ten) betwen the
corrections between the first three quantum states, but I guess this is because
the orbitals behave pretty differently. The fact that the relativistic differences
have nearly a ten-fold difference in energy might indicate that the electrons have
vastly different speeds, which is an interesting result.\\

\end{document}