### 2016-02-24 bioinformatics/latex/a3.tex
blarbl

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Bioinformatics Assignment 2}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Phylogenetic Trees}
For the following set of nucleotides, we compute the number of substitutions between
each sequence and plug it into the Jukes-Cantor distance formula to estimate the
phylogenetic tree for the four species:

\begin{align*}
  \mbox{Human} \hspace{1cm}ACGTGCTGCTAGCTGACTGATCGATCGTACGTCTAG\\
  \mbox{Chimp} \hspace{1cm}ACGTCCTGCAAGCTGACAGATCGATCCTACGTCTAG\\
  \mbox{Mouse} \hspace{1cm}ACGACCAGCTAGGACAGACTTGGATCCTACCTTTAC\\
  \mbox{Rat}   \hspace{1cm}TCGACCAGCTAGGAAAGACTTGCATCCTACCATTAC\\
\end{align*}

\begin{align*}
  \mbox{Substitutions} \hspace{1cm}
  \begin{pmatrix}
    * & 4 & 15 & 18\\
    & * & 13 & 16\\
    & & * & 4\\
    & & & *\\
  \end{pmatrix}
\end{align*}

\end{document}