### 2016-03-13 bioinformatics/latex/a5.tex
crystal files and final BI hw.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{listings}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Bioinformatics Assignment 3}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Sequence Complexity}
We use the Wooton-Federhen complexity score to quantify the complexity of four nucleotide sequences.
The formula is this:

\begin{align*}
  C_{wf} &= \frac{1}{N}\log_4\left(\frac{N!}{n_A!n_C!n_T!n_G!}\right)\\
\end{align*}

Complexity Scores:
\begin{align*}
  &ATATATATATATATATATAT: N=20, n_A=10, n_T=10 \rightarrow C_{wf}
  = \frac{1}{20}\log_4\left(\frac{20!}{10!10!}\right) = 0.43\\
  &GCGCGCGCGCGCGCGCGCGC: N=20, n_G=10, n_C=10 \rightarrow C_{wf}
  = \frac{1}{20}\log_4\left(\frac{20!}{10!10!}\right) = 0.43\\
  &ACGTACGTACGTACGTACGT: N=20, n_A=5, n_T=5 n_G=5, n_C=5 \rightarrow C_{wf}
  = \frac{1}{20}\log_4\left(\frac{20!}{5!5!5!5!}\right) = 0.83\\
  &AAAAAAAAAAAAAAAAAAAA: N=20, n_A=20 \rightarrow C_{wf}
  = \frac{1}{20}\log_4\left(\frac{20!}{20!}\right) = 0\\
\end{align*}

These results are as we expect. The first two sequences are medium complexity, and the one with
an even distribution of the four nucleotides is the highest complexity.\\

\section{Problem Two: }
Here we use the Benjamini-Hochberg procedure for an FDR of 0.01. First, we rank the p-values
ascendingly:

\begin{align*}
  &p_1 \leq p_7 \leq p_2 \leq p_3 \leq p_4 \leq p_5 \leq p_6 \leq p_{10} \leq p_8 \leq p_9\\
  &1\hspace{0.65cm}2\hspace{0.65cm}3\hspace{0.65cm}4\hspace{0.65cm}5\hspace{0.65cm}6\hspace{0.65cm}
  7\hspace{0.65cm}8\hspace{0.65cm}9\hspace{0.65cm}10\\
\end{align*}

We find the maximum significant rank through:

\begin{align*}
  \frac{p_{r*}N}{r*} \leq FDR\\
  \frac{p_{r*}}{r*} \leq 0.001\\
\end{align*}

we find that the $p_1$ and $p_7$ are the only p-values that make this cut, so they are the only
ones we consider significant.\\

\section{Problem 3: }
Do this last

\section{Problem 4: Phylogenetic Tree NN interchange}
Here we use the NNI procedure to come up with new ways of representing the following tree:

\begin{verbatim}
A    B
I----I
C    D
\end{verbatim}

For any four-subtree tree, there are only two ways it can be rearranged. These are the only
two possible rearrangements:

\begin{verbatim}
A    C        A    C
I----I  and   I----I
B    D        D    B
\end{verbatim}

Each tree can be simplified to a pairing between A and one of the other three subtrees, since
you only need one pair to define a tree of size 4. There are three possible pairs involving A,
so three possible trees.\\

\section{Problem Six: ChipSeq}

\end{document}