### 2016-04-15 stat_mech/latex/a1.tex
blablabl

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Statistical Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Adiabatic Ideal Gas}
\textbf{a.)}
First we calculate the final temperature of the gas, given $V_i, V_f$ and $T_i$:

\begin{align*}
  P_iV_i^\gamma &= nRT_iV_i^{\gamma-1}\\
  nRT_iV_i^{\gamma-1} &= nRT_fV_f^{\gamma-1}\\
  T_f &= T_i\left(\frac{V_i}{V_f}\right)^{\gamma-1}
\end{align*}

\textbf{b.)}
If $dS=0$, then $dQ/T=0$ and so $dU = -pdV$. We use this as another route for finding $T_f$ in
terms of $V_i, V_f$ and $T_i$:

\begin{align*}
  dU &= -pdV = C_vdT\\
  -\frac{RT}{V}dV &= C_vdT \hspace{2cm}\mbox{Molar specific heat, so n=1}\\
  -R\int_{V_i}^{V_f}\frac{1}{V}dV &= C_v\int_{T_i}^{T_f}\frac{1}{T}dT\\
  -R\ln\left(\frac{V_f}{V_i}\right) &= C_v\ln\left(\frac{T_f}{T_i}\right)\\
  -\frac{C_p-C_v}{C_v}\ln\left(\frac{V_f}{V_i}\right) &= \ln\left(\frac{T_f}{T_i}\right)\\
  \ln\left(\frac{V_f}{V_i}\right)^{1-\gamma} &= \ln\left(\frac{T_f}{T_i}\right)\\
  \left(\frac{V_f}{V_i}\right)^{1-\gamma} &= \frac{T_f}{T_i}\\
  T_f &= T_i\left(\frac{V_f}{V_i}\right)^{1-\gamma} = T_i\left(\frac{V_i}{V_f}\right)^{\gamma-1}\\
\end{align*}

\section{Problem Two: Sound compression of a liquid}
We use the following variable scheme: $\Delta x$ is the initial length of our 1D
liquid. $\epsilon$ is the change in our 1D liquid length due to rarefaction
on the left side. $\epsilon + \Delta \epsilon$ is the change in length to the right
side.

\begin{align*}
  K_S &= -\frac{1}{V}\left(\frac{dV}{dP}\right)_S
  = \frac{-1}{V}\frac{\Delta V}{\Delta P}& \hspace{2cm}&\mbox{P-P0: displacement from eq pressure}\\
  &= \frac{-1}{A\Delta x}\frac{A\Delta\epsilon}{P-P_0}& \hspace{2cm}&\mbox{A: Area}\\
  \Delta P &= -\frac{1}{K_s}\frac{d\epsilon}{dx}\\
\end{align*}

We then need to find the time derivative of $\epsilon$, which is the force on our
1D cell, or how fast it is growing. The pressure wave is moving in the negative
x direction, so the force on the right of the cell is $P_0 + \Delta P$. On the
left, our cell is being pushed by the normal pressure $P_0$. Thus by Newton's First
Law we see that:\\

\begin{align*}
  m\left(\frac{d^2\epsilon}{dt^2}\right) &= P_0A - (P_0+\Delta P)A = -\Delta PA\\
  \rho A \Delta x \frac{d^2\epsilon}{dt^2} &= -A\frac{dP}{dx}\Delta x\\
  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{dP}{dx}\\
  \rho \frac{d^2\epsilon}{dt^2} &= -\frac{d \Delta P}{dx}\\
  \rho \frac{d^2}{dt^2}\frac{d\epsilon}{dx} &= -\frac{d^2P}{dx^2}\\
  \rho \frac{d^2}{dt^2}\left(-PK_s\right) &= -\frac{d^2P}{dx^2}\\
  \rho K_s \frac{d^2P}{dt^2}\left(P\right) &= \frac{d^2P}{dx^2}\\
  \frac{d^2P}{dt^2} &= \frac{1}{\rho K_s }\frac{d^2P}{dx^2}\\
\end{align*}

Solution inspired by several people from SPS.\\

\section{Problem Three: Solid $\Delta T, \Delta P$ relation}

\begin{align*}
  dS &= \left(\frac{dS}{dT}_P\right)dT + \left(\frac{dS}{dP}_T\right)dP = 0\\
  \left(\frac{dS}{dT}_P\right)dT &= -\left(\frac{dS}{dP}_T\right)dP\\
  \frac{C_P}{T}dT &= \left(\frac{dV}{dT}_P\right)dP\\
  \frac{C_P}{T}dT &= V\alpha dP\\
  \frac{C_P}{TV\alpha} &= -\left(\frac{dP}{dT}_S\right)\\
  \frac{C_P}{TV\alpha} &= \frac{\Delta P}{\Delta T}\\
  \Delta T &= \frac{TV\alpha}{C_P} = \frac{\alpha T}{\rho c_p} \hspace{2cm}
  c_p \mbox{ is the volumetric specific heat}
\end{align*}

\section{Problem Four: Gas Engine}
Here we use the ideal gas property $dU = C_v$ to express all changes in internal energy
as changes in temperature. We also use the adiabatic $PV^\gamma = k$ relation to convert
between variables during the insulated work-doing steps:\\

We first find the work done:
\begin{align*}
  W_{ab} &= -dU \hspace{2cm}\mbox{Q = 0, First law of Thermo}\\
  &= -C_vdT \hspace{2cm}dU_{ideal} = CvdT\\
  &= -C_v\left(T_b-T_a\right)\\
  &= -C_vT_b\left(1-\frac{T_a}{T_b}\right)\\
  &= -C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
             \hspace{2cm}P_aV_2^\gamma = P_bV_1^\gamma \rightarrow
             \frac{T_a}{T_b} = \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
  W_{cd} &= C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
\end{align*}

Then we find the heat added during the heating step:

\begin{align*}
  dQ_1 &= dU_V = C_v\left(T_c-T_b\right)\\
\end{align*}

Now we find the efficiency:

\begin{align*}
  \eta &= \frac{W}{Q_1} = \frac{C_vT_c\left(1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)
    - C_vT_b\left(1-\left(\frac{V_2}{V_1}\right)^{\gamma-1}\right)}{C_v\left(T_c-T_b\right)}\\
  &= 1 - \left(\frac{V_2}{V_1}\right)^{\gamma-1}\\
\end{align*}

\end{document}