### 2016-05-16 stat_mech/latex/a3.tex
a3 statmech

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{marvosym} % Required for symbols in the colored box
\usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Statistical Mechanics Assignment Three}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem One: Minimizing a 4-edged box}
\textbf{a.) Calculus}

\begin{align*}
  A &= 2xy + 2yz + xz\\
  V = xyz\\
  \grad V &= \left(\frac{d}{dx} + \frac{d}{dy} + \frac{d}{dz} \right)xyz = 0\\
  xy + yz + xz = 0\\
  A_{max} &= 2xy + 2yz - xy - yz = xy + yz\\

\end{align*}

\section{Problem Two: Two-system entropy}
\textbf{a.)} The probability of two unrelated events occuring is the product of the
two individual probabilities. Because the two systems are weakly interacting,
they are ``unrelated'' and so $P_{rs} = P_rP_s$.\\

\textbf{b.)}
\begin{align*}
  S &= -k_B\sum_r \sum_s P_{rs} \ln\left(P_{rs}\right)\\
  &= -k_B\sum_r \sum_s P_rP_s \ln\left(P_rP_s\right)\\
  &= -k_B\sum_r \sum_s P_rP_s\left(\ln P_r + \ln P_s\right)\\
\end{align*}

\end{document}