### 2016-05-24 stat_mech/latex/a4.tex
Add statmech 4

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{graphicx}
\usepackage{amsmath}
%% \usepackage{marvosym} % Required for symbols in the colored box
%% \usepackage{ifsym} % Required for symbols in the colored box

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

\allowdisplaybreaks

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Statistical Mechanics Assignment Four}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem Two: Two-Dimensional Gas}
To tackle this problem, we start by finding the 1-D ideal gas partition function, use it to get the
2-D partition function, use this new partition function to find the energy, then use the energy
to find the heat capacity of a 2-D ideal gas.\\

1-D Partition function:
\begin{align*}
  Z_x &= \int_{-\infty}^{\infty} e^{-mv_x^2/2kT} dv_x\\
  &= \sqrt{\frac{2kT\pi}{m}}\\
\end{align*}

2-D Partition function:
\begin{align*}
  Z &= Z_xZ_y = Z_x^2 = \frac{2kT\pi}{m}\\
\end{align*}

We then find the energy of the gas:

\begin{align*}
  <E> &= N\sum_{i} \epsilon_i p_i\\
  &= \frac{N}{Z}int_{-\infty}^{\infty}int_{-\infty}^{\infty}
  \frac{m(v_x^2+v_y^2)}{2}e^{-m(v_x^2+v_y^2)/2kT} dv_xdv_y\\
  &= \frac{Nm}{2kT\pi}
  \int_{0}^{\infty}\int_0^{2\pi}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
  &= \frac{Nm}{2kT\pi}
  \int_{0}^{\infty}\frac{mv_r^2}{2}e^{-mv_r^2/2kT}}v\sin\theta dvd\theta\\
\end{align*}

\end{document}