### 2016-10-19 em/latex/HW2.tex
Add EM3 work so far.

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Elegromagnetism Assignment Two}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\newcommand{\bra}[1]{\big<#1\big|}
\newcommand{\ket}[1]{\big|#1\big>}
\newcommand{\braket}[2]{\big<#1\big|#2\big>}

\begin{document}

\maketitle{}

\section{Problem One}
To find the bound charges on this cross section, we use the following equations:

\begin{align*}
  \rho_b &= -\nabla \vec{P} = 0\hspace{2cm} \mbox{P is constant}\\
  \sigma_b &= \vec{P} \cdot \vec{n}\\
\end{align*}

The top two faces each make a $60^\circ$ angle with $\vec{P}$, and the bottom one is antiparallel. Thus we see:

\begin{align*}
  \sigma_{b,top} &= \vec{P} \cdot \hat{n} = P\cos\left(60^\circ\right) = \frac P2\\
  \sigma_{b,bot} &= \vec{P} \cdot \hat{n} = P\cos\left(180^\circ\right) = -P\\
\end{align*}

To find total charge we just sum these densities over the surface area of the prism. We notice that there are two top faces and one
bottom face, thus the two surface charges cancel out and the net charge on the object is \textbf{zero}. We expect this for a dielectric with no free charge.\\

We then find the far field by treating the triangle as a point dipole extending infinitely through the z axis:

\begin{align*}
  \vec{E}(\vec{r}, \theta) &= \int_\infty^\infty \frac{\vec{P}\cdot\vec{r}}{\sqrt{z^2+r^2}}dz
  = P\cos(\theta)\int_\infty^\infty \frac{\sqrt{z^2+r^2}}{\sqrt{z^2+r^2}^3}dz
  = P\cos(\theta)\int_\infty^\infty \frac{1}{z^2+r^2}dz\\
  &= \frac{P\cos(\theta)\pi}{r}
\end{align*}

This result almost makes sense. Normally a dipole falls off by $\frac{1}{r^3}$, but this one is generated by an infinite line, so it
falls off lots slower.\\

\section{Problem Two}
The example solved in Griffiths Ex 4.7 can be generalized to a two-dielectrics-in-all-space system by simply swapping the $\epsilon_0$ for vacuum with the dielectric constant of the outer dielectric, $\epsilon_r^2$. Thus the solution is identical to the book's plus this swap:

\textbf{a.)}\\
\begin{align*}
	\epsilon &= \epsilon_r^1/\epsilon_r^2\\
	V_{in}(r,\theta) &= -\frac{3E_0r\cos\theta}{\epsilon+2}\\
	E_{in} &= \frac{3}{\epsilon+2}E_0\hat{z}\\
        V_{out}(r,\theta) &= -E_0r\cos\theta + \frac{\epsilon-1}{\left(\epsilon+2\right)r^2}\cos(\theta)\\
        E_{out}(r,\theta) &= \left(E_0 + 2\frac{\epsilon-1}{\left(\epsilon+2\right)r^3\right)\cos\left(\theta\right)\hat{r}
        + \left(-E_0 + \frac{\epsilon-1}{\left(\epsilon+2\right)r^3}\right)\cos\left(\theta\right)\hat{\theta}\\
\end{align*}\\

\textbf{b.)}\\
For linear dielectrics we can find $\sigma_b$ from the P vector, which comes from from $\vec{E}$:

\begin{align*}
  \vec{P}_{in} &= \left(\epsilon-\epsilon_0\right)\vec{E}_{in}\\
  &= 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\hat{z}\\
  \sigma_b &= \vec{P} \cdot \hat{r} = 3E_0\frac{\epsilon-\epsilon_0}{\epsilon+2}\cos(\theta)
\end{align*}

This charge is positive at the top of the sphere and negative at the bottom, as expected for a bound charge.\\

\textbf{c.)}\\
We now show that the tangential E-fields are continuous at the surface. We note that there are no discontinuities in either dot product for any value of $\theta$:

\begin{align*}
  E_{in}(R)\cdot\hat{\theta} &= \frac{3}{\epsilon+2}E_0\cos(\theta)\\
  E_{out}(R)\cdot\hat{\theta} &= \frac{\epsilon-1}{\epsilon+2}E_0\sin\theta
\end{align*}

\textbf{d.)}
$D_\theta$ can be arrived at from $E_{\theta}$ by multiplying by a factor of $\epsilon_r^1$ or $\epsilon_r^2$.

\begin{align*}
  D_{in}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta)\\
  D_{out}(R) &= \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
  D_{in}(R) - D_{out}(R) &= \frac{3\epsilon_r^1}{\epsilon+2}E_0\cos(\theta) - \frac{\epsilon_r^1-\epsilon_r^2}{\epsilon+2}E_0\sin\theta\\
  &= \frac{E_0}{\epsilon+2}\left(3\epsilon_r^1\cos(\theta) - \left(\epsilon_r^1 - \epsilon_r^2\right)\sin(\theta)\right)
\end{align*}

\end{document}