### 2017-02-12 311/latex/HW3.tex
HW3.tex

\documentclass[10pt]{article} % Font size - 10pt, 11pt or 12pt

\nonstopmode

\usepackage[hmargin=1.25cm, vmargin=1.5cm]{geometry} % Document margins

\usepackage{amsmath}

\usepackage[usenames,dvipsnames]{xcolor} % Allows the definition of hex colors

% Fonts and tweaks for XeLaTeX
\usepackage{fontspec,xltxtra,xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmonofont[Scale=MatchLowercase]{Andale Mono}

% Colors for links, text and headings
\usepackage{hyperref}
\definecolor{linkcolor}{HTML}{506266} % Blue-gray color for links
\definecolor{shade}{HTML}{F5DD9D} % Peach color for the contact information box
\definecolor{text1}{HTML}{2b2b2b} % Main document font color, off-black
\definecolor{headings}{HTML}{701112} % Dark red color for headings
% Other color palettes: shade=B9D7D9 and linkcolor=A40000; shade=D4D7FE and linkcolor=FF0080

\hypersetup{colorlinks,breaklinks, urlcolor=linkcolor, linkcolor=linkcolor} % Set up links and colors

\usepackage{fancyhdr}
\usepackage{amssymb}
\pagestyle{fancy}
\fancyhf{}
% Headers and footers can be added with the \lhead{} \rhead{} \lfoot{} \rfoot{} commands
% Example footer:
%\rfoot{\color{headings} {\sffamily Last update: \today}. Typeset with Xe\LaTeX}

\renewcommand{\headrulewidth}{0pt} % Get rid of the default rule in the header

\usepackage{titlesec} % Allows creating custom \section's

% Format of the section titles
\titleformat{\section}{\color{headings}
\scshape\Large\raggedright}{}{0em}{}[\color{black}\titlerule]

\title{Advanced Calc Homework 3}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

\begin{document}

\maketitle{}

\section{Problem 2.2.2}

\section{Problem 2.4.4}
\textbf{a.} $\{\frac{1}{n}\}$\\
Peak at $n=1$. $\forall m \in \mathbb{N}$, $\frac{1}{n} \geq \frac{1}{n+m}$. Thus $a_n \geq a_{n+m}$.\\

\textbf{b.} $\{\left(01\right)^n\}$\\
Peak at even $n$s. For any even $n$, $a_n = 1$ and $\forall m \in \mathbb{N}$, $a_{n+m} = 1$ or $-1$. Thus, $a_n \geq a_{n+m}$.\\

\textbf{c.} $\{(-1)^nn\}$\\
No peaks. $\forall n \in \mathbb{N}$, if $n$ is even then $a_n = n < n+2 = a_{n+2}$. If $n$ is odd, then $a_n = -n < n+1 = a_{n+1}$.\\

\textbf{d.} $\{\frac{(-1)^n}{n}\}$\\
Peak at $n=2$. $\forall m \in \mathbb{N}$, if $m$ is odd then $a_{2+m} < 0 < a_{2}$. If $m$ is even then $a_{n+m} = \frac{1}{2+m} < \frac{1}{2}$.\\

\section{Problem 2.4.7}

\end{document}