### 2018-09-29 test/report.tex
bla

\documentclass[a4paper, 12pt]{article}
\usepackage{fullpage} % changes the margin

\begin{document}
\noindent
\huge\textbf{Project report}
\normalsize

\section*{Introduction}

\section*{Objectives}

\begin{itemize}
\item Create a biologically realistic leaky integrate-and-fire (LIF) model of a cortical section which can reproduce single-neuron details from 2PI \cite{bellaypaper}, such as firing rate distribution and quiescent time, correlations (average cross-correlation, distance-dependent correlation and autocorrelation), branching parameter, and time-binned avalanche dynamics

\item Use biophysical hierarchy of neural connections \cite{rubinov} and represent the main pyramidal and interneuron types within cortex using realistic parameters \cite{objectworkingmemory}

\item Reproduce the theoretical synchronous-to-asynchronous phase transition \cite{munozlg} using a synaptic resources parameter tuned to avoid the bistable up-down regime and non-critical (neutral) avalanches. Other models have found a continuous (presumably not up/down) phase transition \cite{rubinov}, so it is likely we will be able to find one as well. Other models have remained within the bistable regime and thus see only an up/down transition with neutral avalanches \cite{neutraltheory}.

\item Use the 'exact' solution method \cite{exactsolution} to quickly simulate large networks, possibly making the model fast enough to explore the parameter space to find different network modes (eg oscillation, up/down bistability \cite{munozlg}), or to optimize via gradient descent

\item schizophrenia - something something either simulate PCP problems biophysically and try to reproduce the PCP phenotype OR manually just increase the repeat-firing of cortical neurons

\item Possibly introduce a spatial component to neurons to estimate LFP electrode measurements from the LIF network, allowing us to see if the model can be made to reproduce coherence potentials
\end{itemize}

\section*{Methods}

\subsection*{Modular hierarchy}
Recent models for self-organized criticality have used neural models with hundreds to thousands of tightly-connected neurons \cite{munozlg, rubinov}. We will use a similar scheme to compare with these results.\\

\subsection*{Connectivity}
\subsection*{Plasticity \& synaptic resources}
\subsection*{Voltage equations}

\section*{TODO}
\begin{itemize}
\item Find papers on LIF cortical organization - how should we connect the modules? Should we do modules? What should they look like?
\item Think about what sorts of plasticity rules would accomodate the Munoz 'synaptic resources' property - a way to make modules have propensity to do module spikes if sufficiently excited, and also a refractory period where the whole module must recover
\item Are there any 'more' biophysical models which could be used to get more interesting results? Anything which would reproduce interesting behavior which LIF cannot, eg realistic connectivity between pyramidal and interneurons?
\end{itemize}

\textbf{coherence potentials subproject}
\begin{itemize}
\item Could CPs be found in 2P data?
\item Are there any patterns in MEA CP traces suggestive of how they arise?
\item Could large-scale synchronous dynamics in the brain, eg oscillations or avalanches, \textit{create} hopfield networks in different places? Seems unlikely since CPs are \textit{really} similar
\item Are CPs clustered in space, or are they a small set of repeating patterns which the brain likes to make in different places? Does injecting a CP into a network make it more likely for it to be seen at a later time?
\item Are there any processes which could move a conceivably-CP-creating strong hopfield network \textit{spatially} through a network?
\end{itemize}

\bibliographystyle{plain}
\bibliography{report}

\end{document}